---
title: "Parameter estimation"
author: "Richard J Beck"
date: "6/13/2022"
output: html_document
---


This code generates the output for figure 5 and the associated supplementary figures
  
Data used by this file can be generated using optim_script.R optim_script_ko.R and run_amphs.R scripts contained in code/utils/  

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="path/to/LAG_TIM_BIOMARKERS/")
```

```{r}
library(pbapply)
library(parallel)
library(deSolve)
library(ggplot2)
source("code/utils/optim_funcs.R")
source("code/utils/plotting_funcs.R")


## the following just splits up the amphs chain into separate files to get round githubs file size limit
pack_amphs_chain <- function(chain_path){
  x <- readRDS(chain_path)
  lapply(1:length(x), function(i){
    saveRDS(x[[i]],paste0("data/RESULTS/unpacked_chain/",i,".Rds"))
  })
}

unpack_amphs_chain <- function(unpacked_chain_path){
  fnames <- paste0(unpacked_chain_path,list.files(unpacked_chain_path))
  x <- lapply(fnames,readRDS)
  return(x)
}

chain_path <- "data/RESULTS/all_pars_wider_chain.Rds"
#pack_amphs_chain(chain_path) I ran this to initially split the chain
unpacked_chain_path <- "data/RESULTS/unpacked_chain/"
x <- unpack_amphs_chain(chain_path) ## this will unsplit the chain
saveRDS(x, chain_path)
```

Estimate the SD for the inhibitors:
Approach is to take a list of housekeeping genes (https://doi.org/10.1093/nar/gkaa609) (HRT Atlas v1.0, www.housekeeping.unicamp.br) (Most stable mouse transcripts) and calculate the relationship between their mean and sd in our data.

We get sd = 0.054+0.29*mean
```{r}
df <- read.csv("data/GENE_EXPRESSION/PROCESSED/all_gProcessedSignal_BGfiltered.csv")

x <- df[,7:ncol(df)]

q75 <- apply(x,2,quantile)["75%",]
x <- x/q75

hkg <- read.csv("data/GENE_EXPRESSION/MostStable_Mouse.csv",sep=";")

x <- x[df$GeneName%in%hkg$Gene.name|df$SystematicName%in%hkg$transcript_id,]
sdx <- apply(x,1,sd)
mx <- apply(x,1,mean)

x$GeneName <- sapply(1:nrow(x), as.character)
df <- reshape2::melt(x,id.vars="GeneName")
df$condition <- sapply(as.character(df$variable), function(i) unlist(strsplit(i,split="_"))[1])

p <- ggplot(df,aes(x=GeneName,y=value,color=condition))+
  geom_point()+
  scale_x_discrete("housekeeping genes")+
  scale_y_log10("expression value")+
  theme(axis.text.x = element_blank())
p

mod <- lm(sdx~mx)
mean_expression <- mx
stdev_expression <- sdx
plot(mean_expression,stdev_expression)
lines(mx,predict(mod,list(mx)))

print(mod)

```


check fits


```{r}
base_plot_size=8
margin_size <- 15

x <- readRDS("data/RESULTS/opt_out_wider_bounds.Rds")
pars <- do.call(rbind,lapply(x, function(xi) xi$par))
pz <- readRDS("data/RESULTS/all_pars_wider_chain.Rds")[[1]]$par
pz <- pz[round(nrow(pz)/2):nrow(pz),1:3]
pz <- data.frame(exp(pz))
colnames(pz) <- c("k[l]","k_[t]","k_[p]")
pz$id <- 1:nrow(pz)
pz <- reshape2::melt(pz,id.vars=c("id"))

p5f <- ggplot(pz,aes(x=value))+
  facet_wrap(~variable,scales="free",labeller=label_parsed)+
  geom_histogram(bins = 20)+
  #geom_density()+
  scale_x_log10("parameter value",guide = guide_axis(check.overlap = TRUE))+
  theme_classic(base_size=base_plot_size)
p5f

v <- sapply(x, function(xi) xi$value)
pars <- pars[order(v),]
v <- v[order(v)]


x <- readRDS("data/RESULTS/ko_k_LAG_k_TIM_opt_out.Rds")
px <- do.call(rbind,lapply(x, function(xi) xi$par))
vx <- sapply(x, function(xi) xi$value)
px <- px[vx==min(vx),]
px[6] <- px[5]
vx <- vx[order(vx)]

pb <- pars[1,]

x1 <- load_fit("TIM\ndominant",pars[1,])
x2 <- load_fit("LAG\ndominant",pars[6,])


a <- rbind(load_activity("TIM\ndominant",pars[1,]),
           load_activity("LAG\ndominant",pars[6,]))

##contribution to CTL exhaustedness
p5e <- ggplot(a,aes(x=time,y=value,color=variable))+
  facet_grid(cols=vars(id))+
  geom_line()+
  scale_y_continuous("relative contribution\nto exhaustion")+
  scale_x_continuous("days after CTL transfer")+
  scale_color_discrete("",labels=c("PD1/PDL1","Lag3","Tim3"))+
  theme_classic(base_size=base_plot_size)
p5e

## CTL activity
p5d <- ggplot(a,aes(x=time,y=E_activ,color=id,linetype=id,size=id))+
  geom_line()+
  scale_y_continuous("CTL activity")+
  scale_x_continuous("days after CTL transfer")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(1,1.3))+
  theme_classic(base_size=base_plot_size)
p5d


df <- rbind(x1$df,x2$df)
df_inh <- rbind(x1$df_inh,x2$df_inh)


d0 <- setup_opt(path = "data/PAR_EST/")
dat <- reshape_data(d0$dat)
dat1 <- dat[dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
dat <- dat[!dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]

dfv <- data.frame(logL = -v, id=1:length(v))

dfp <- data.frame(exp(pars[1:11,]))
colnames(dfp) <- parnames
dfp$id <- 1:nrow(dfp)
dfp$logL <- -v[1:nrow(dfp)]

dfp <- reshape2::melt(dfp,id.vars=c("id","logL"))

pp <- ggplot(dfp,aes(x=variable,y=value))+
  geom_jitter(height=0,width=0.1)+
  scale_y_log10("parameter value")+
  scale_x_discrete("")+
  theme_bw()
pp


pv1 <- ggplot(dfv,aes(x=id,y=-logL))+
  geom_point()+
  scale_x_log10("number of repeats")+
  scale_y_log10("log likelihood")+
  theme_bw()
pv1

pv2 <- ggplot(head(dfv,21),aes(x=id,y=logL))+
  geom_point()+
  scale_x_continuous("number of repeats")+
  scale_y_continuous("log likelihood")+
  theme_bw()
pv2

id_renamr <- c(E="CTLs",g="g (day^-1)",gr="S-G2-M:G1",IFN="Ifng",
  LAG3="Lag3",TIM3="Tim3",PD1="PD-1",PDL1="PD-L1")

df$variable <- id_renamr[as.character(df$variable)]
dat$variable <- id_renamr[as.character(dat$variable)]

p5b <- ggplot(df)+
  facet_grid(rows=vars(variable),scales="free",labeller = label_parsed)+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id),show.legend=F)+
  geom_point(data=dat,aes(x=day,y=mean))+
  geom_errorbar(data=dat,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
  scale_x_continuous("days")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(0.8,1))+
  theme_classic(base_size=base_plot_size)+
  theme(plot.margin = margin(4,4,4,8, "pt"))
p5b

df_inh$variable <- id_renamr[as.character(df_inh$variable)]
dat1$variable <- id_renamr[as.character(dat1$variable)]

p5c <- ggplot(df_inh)+
  facet_grid(rows=vars(variable),scales="free")+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id))+
  geom_point(data=dat1,aes(x=day,y=mean))+
  geom_errorbar(data=dat1,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
    scale_x_continuous("days")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(0.8,1))+
  theme_classic(base_size=base_plot_size)
p5c


```

```{r}
base_plot_size=10
margin_size <- 15
x <- readRDS("data/RESULTS/all_pars_wider_chain.Rds")[[1]]
par <- data.frame(exp(x$par))
par <- par[round(nrow(par)/2):nrow(par),]

x <- par

par$id <- 1:nrow(par)

par <- reshape2::melt(par,id.vars="id")

levels(par$variable) <- c(k_LAG="k[l]",k_TIM="k[t]",k_PD="k[p]",k_inf="s[0]",
                          k_exp="s[e]",de="d[e]",d_lag="d[l]",d_tim="d[t]",
                          d_pd1="d[p]",d_pdl1="d[pl]",d_ifn="d[i]",
                          k_ifn="k[i]",ke="k[e]")[levels(par$variable)]
p <- ggplot(par,aes(x=value))+
  facet_wrap(~variable,scales="free",labeller=label_parsed)+
  geom_histogram(bins = 20)+
  #geom_density()+
  scale_x_log10("parameter value",guide = guide_axis(check.overlap = TRUE))+
  theme_classic()
p

p2 <- ggplot(x,aes(x=k_LAG,y=k_TIM))+
  #geom_point(alpha=0.1)+
  geom_bin_2d()+
  scale_x_log10(expression(Lag3~strength~(k[l])),breaks=c(0.001,0.1,10))+
  scale_y_log10(expression(atop("Tim3 strength",(k[t]))),breaks=c(0.001,0.1,10))+
  theme_classic(base_size=base_plot_size)+
  scale_fill_viridis_c("MCMC\nsamples")+
  theme(plot.margin = margin(3,3,3,3, "pt"),legend.key.size = unit(10,"pt"))
p2

p <- ggplot(x,aes(x=k_LAG,y=d_lag))+
  #geom_point(alpha=0.1)+
  geom_bin_2d()+
  scale_x_log10(expression(Lag3~inhibitor~strength~k[l]))+
  scale_y_log10(expression(d[l]))+
  theme_classic()+
  scale_fill_viridis_c("MCMC\nsamples")
p
p <- ggplot(x,aes(x=k_TIM,y=d_tim))+
  #geom_point(alpha=0.1)+
  geom_bin_2d()+
  scale_x_log10(expression(k[t]))+
  scale_y_log10(expression(d[t]))+
  theme_classic()+
  scale_fill_viridis_c()
p

p <- ggplot(x,aes(x=ke,y=d_pd1))+
  #geom_point(alpha=0.1)+
  geom_bin_2d()+
  scale_x_log10(expression(k[e]))+
  scale_y_log10(expression(d[p]))+
  theme_classic()+
  scale_fill_viridis_c()
p

p1 <- ggplot(par[par$variable%in%c("k[l]","k[t]","k[p]"),],aes(x=value))+
  facet_wrap(~variable,ncol=1,labeller=label_parsed)+
  geom_histogram(bins = 20)+
  #geom_density()+
  scale_x_log10("",guide = guide_axis(check.overlap = TRUE))+
  scale_y_continuous("MCMC samples")+
  theme_classic(base_size=base_plot_size)+
  theme(plot.margin = margin(3,3,3,5, "pt"))
p1


aic <- readRDS("data/RESULTS/aic_dataset.Rds")

paic <- ggplot(aic,aes(x=aic,y=id))+
  geom_col()+
  scale_y_discrete("")+
  scale_x_continuous("AIC")+
  theme_classic(base_size = base_plot_size)+
  theme(plot.margin = margin(3,3,3,5, "pt"))
paic




```

Various knockout experiments

```{r}

tmp <- function(fixpars){
  parnames <- c("k_LAG","k_TIM","k_PD",
              "k_inf","k_exp","de",
              "d_lag","d_tim","d_pd1","d_pdl1", "d_ifn",
              "k_ifn","ke")
  
  p1 <- pars[1,]
  p2 <- pars[6,]
  names(p1) <- parnames
  names(p2) <- parnames
  
  if(!is.null(fixpars)){
    p1 <- p1[!names(p1)%in%names(fixpars)]
    p2 <- p2[!names(p2)%in%names(fixpars)]
  }
  
  x1 <- load_fit("TIM\ndominant",p1,fixpars=fixpars)
  x2 <- load_fit("LAG\ndominant",p2,fixpars=fixpars)

  df <- rbind(x1$df,x2$df)
  dfi <- rbind(x1$df_inh,x2$df_inh)
  
  if(is.null(fixpars)) fixpars=c(none=0)
  df$fixpars <- names(fixpars)
  dfi$fixpars <- names(fixpars)
  list(df=df,dfi=dfi)
}

base_plot_size=10


x <- readRDS("data/RESULTS/opt_out_wider_bounds.Rds")


pars <- do.call(rbind,lapply(x, function(xi) xi$par))
v <- sapply(x, function(xi) xi$value)
pars <- pars[order(v),]
v <- v[order(v)]


x <- readRDS("data/RESULTS/ko_k_LAG_k_TIM_opt_out.Rds")
px <- do.call(rbind,lapply(x, function(xi) xi$par))
vx <- sapply(x, function(xi) xi$value)
px <- px[vx==min(vx),]
px[6] <- px[5]
vx <- vx[order(vx)]

pb <- pars[1,]

x0 <- tmp(NULL)
x1 <- tmp(fixpars=c(k_LAG=0))
x2 <- tmp(fixpars=c(k_TIM=0))
x3 <- tmp(fixpars=c(k_PD=0))

df <- rbind(x0$df,x1$df,x2$df,x3$df)
dfi <- rbind(x0$dfi,x1$dfi,x2$dfi,x3$dfi)
d0 <- setup_opt(path = "data/PAR_EST/")
dat <- reshape_data(d0$dat)
dat1 <- dat[dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
dat <- dat[!dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]

fixpar_renamr <- c(k_LAG="k[l]==0",
                   k_TIM="k[t]==0",
                   k_PD="k[p]==0",
                   none="full \nmodel")

df$fixpars <- fixpar_renamr[df$fixpars]
dfi$fixpars <- fixpar_renamr[dfi$fixpars]

id_renamr <- c(E="CTLs",g="g (day^-1)",gr="S-G2-M:G1",IFN="Ifng",
  LAG3="Lag3",TIM3="Tim3",PD1="PD-1",PDL1="PD-L1")

df$variable <- id_renamr[as.character(df$variable)]
dat$variable <- id_renamr[as.character(dat$variable)]
dfi$variable <- id_renamr[as.character(dfi$variable)]
dat1$variable <- id_renamr[as.character(dat1$variable)]

s1a <- ggplot(df)+
  facet_grid(rows=vars(variable),cols=vars(fixpars),scales="free",labeller = "label_parsed")+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id))+
  geom_point(data=dat,aes(x=day,y=mean))+
  geom_errorbar(data=dat,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
    scale_x_continuous("days")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(0.8,1))+
  theme_classic(base_size=base_plot_size)
s1a

s2a <- ggplot(dfi)+
  facet_grid(rows=vars(variable),cols=vars(fixpars),scales="free",labeller = "label_parsed")+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id))+
  geom_point(data=dat1,aes(x=day,y=mean))+
  geom_errorbar(data=dat1,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
    scale_x_continuous("days")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(0.8,1))+
  theme_classic(base_size=base_plot_size)
s2a

```

Re-running of simulations to determine importance of killing & antiproliferative:
```{r}

tmp <- function(id){
  
  ix <- 1
  if(id=="LAG\ndominant") ix <- 8
  
  x0 <- run_model(pars[ix,])
  x0$id <- "full\nmodel"
  x1 <- run_model(pars[ix,],fixpars=c(ke=0))
  x1$id <- "only\nantiprolif."
  x2 <- run_model(pars[ix,],fixpars=c(k_ifn=0))
  x2$id <- "only\nkilling"

  df <- rbind(x0,x1,x2)
  df <- df[,c("time","SG2M","G1","id")]
  df$total <- df$G1+df$SG2M
  df <- reshape2::melt(df,id.vars=c("time","id"))
  df$id2 <- id
  return(df)
}

base_plot_size=10


x <- readRDS("data/RESULTS/opt_out_wider_bounds.Rds")


pars <- do.call(rbind,lapply(x, function(xi) xi$par))
v <- sapply(x, function(xi) xi$value)
pars <- pars[order(v),]
v <- v[order(v)]

a1 <- load_activity("TIM\ndominant",pars[1,])
a1$ke <- pars[1,ncol(pars)]
a2 <- load_activity("LAG\ndominant",pars[6,])
a2$ke <- pars[6,ncol(pars)]

a <- rbind(a1,a2)
a$killing <- a$E_activ*exp(a$ke)

s1a <- ggplot(a,aes(x=time,y=killing,color=id,linetype=id,size=id))+
  geom_line()+
  scale_y_continuous(expression(killing~rate~(CTL^-1~day^-1)))+
  scale_x_continuous("day after CTL transfer")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(0.8,1))+
  theme_classic(base_size = base_plot_size)
s1a

df <- rbind(tmp("TIM\ndominant"),tmp("LAG\ndominant"))


s1c <- ggplot(df,aes(x=time,y=value,color=variable))+
  facet_grid(rows=vars(id),cols=vars(id2))+
  geom_line()+
  scale_y_log10("tumour cells")+
  scale_color_manual("",values=c("green","red","blue"))+
  scale_x_continuous("day after CTL transfer")+
  theme_classic(base_size = base_plot_size)+
  theme(legend.position = "top")
s1c

s1b <- ggplot(df[df$variable=="total",],aes(x=time,y=value,color=id,linetype=id))+
  facet_grid(rows=vars(id2))+
  geom_line()+
  scale_y_log10("tumour cells")+
  scale_x_continuous("day after CTL transfer")+
  scale_color_discrete("")+
  scale_linetype_manual("",values=c("longdash","dotted","solid"))+
  theme_classic(base_size = base_plot_size)
s1b




```

Practical identifiability



```{r}

base_plot_size=10
margin_size <- 15
x <- readRDS("data/RESULTS/all_pars_wider_chain.Rds")[[1]]
par <- data.frame(exp(x$par))
par <- par[round(nrow(par)/2):nrow(par),]

x <- par

par$id <- 1:nrow(par)

par <- reshape2::melt(par,id.vars="id")

levels(par$variable) <- c(k_LAG="k[l]",k_TIM="k[t]",k_PD="k[p]",k_inf="s[0]",
                          k_exp="s[e]",de="d[e]",d_lag="d[l]",d_tim="d[t]",
                          d_pd1="d[p]",d_pdl1="d[pl]",d_ifn="d[i]",
                          k_ifn="k[i]",ke="k[e]")[levels(par$variable)]


### Construct df to get confidence intervals

parnames <- c("k_LAG","k_TIM","k_PD",
                "k_inf","k_exp","de",
                "d_lag","d_tim","d_pd1","d_pdl1", "d_ifn",
                "k_ifn","ke")

df <- data.frame(variable=parnames,identifiable=c("r","r","r","r","y","y",
                                         "n","n","n","n","y","y","n"))

q <- do.call(rbind,lapply(unique(par$variable), function(v){
  quantile(par$value[par$variable==v],probs=c(0,0.01,0.99,1))
}))

q <- data.frame(do.call(rbind,lapply(1:nrow(q), function(i){
  if(df$identifiable[i]=="r") return(q[i,c(1,3)])
  if(df$identifiable[i]=="l") return(q[i,c(2,4)])
  if(df$identifiable[i]=="y") return(q[i,c(2,3)])
  if(df$identifiable[i]=="n") return(q[i,c(1,4)])
})))

colnames(q) <- c("lower","upper")

df <- cbind(df,q)
df$y <- 2000
df$variable <- c(k_LAG="k[l]",k_TIM="k[t]",k_PD="k[p]",k_inf="s[0]",
                          k_exp="s[e]",de="d[e]",d_lag="d[l]",d_tim="d[t]",
                          d_pd1="d[p]",d_pdl1="d[pl]",d_ifn="d[i]",
                          k_ifn="k[i]",ke="k[e]")[df$variable]
df2 <- df
df <- reshape2::melt(df,id.vars=c("variable","identifiable","y"))
colnames(df)[4]<-"bound"
df <- df[df$identifiable!="n",]



df <- df[!(df$identifiable=="r"&df$bound=="lower"),]
df <- df[!(df$identifiable=="l"&df$bound=="upper"),]
p <- ggplot(par)+
  facet_wrap(~variable,nrow=3,scales="free",labeller=label_parsed)+
  geom_histogram(bins = 20,aes(x=value))+
  geom_vline(data=df,aes(xintercept=value),color="red")+
  scale_x_log10("parameter value",guide = guide_axis(check.overlap = TRUE))+
  scale_y_continuous("MCMC samples")+
  theme_classic()
p


```

```{r}
base_plot_size=10

x <- readRDS("data/RESULTS/opt_out_wider_bounds.Rds")
pars <- do.call(rbind,lapply(x, function(xi) xi$par))
v <- sapply(x, function(xi) xi$value)
pars <- pars[v==min(v),]

names(pars)<- c("k_LAG","k_TIM","k_PD",
                "k_inf","k_exp","de",
                "d_lag","d_tim","d_pd1",
                "d_pdl1", "d_ifn",
                "k_ifn","ke")

y <- rbind(sweep_par(pb = pars,sweep = "d_pd1"),
           sweep_par(pb = pars,sweep = "d_pdl1"),
           sweep_par(pb = pars,sweep = "d_lag"),
           sweep_par(pb = pars,sweep = "d_tim"))

yrenamr <- c(d_pd1="d[p]",d_pdl1="d[pl]",d_lag="d[l]",d_tim="d[t]")

y$swept <- yrenamr[y$swept]
#y$swept <- sapply(y$swept,as.expression)

s5a <- ggplot(y,aes(x=exp(parval),y=ll,color=swept))+
  geom_line()+
  scale_y_continuous("negative logL",limits=c(2.8,10))+
  scale_x_log10("checkpoint disappearance rate")+
  scale_color_discrete("",labels=c(expression(d[l]),
                                   expression(d[p]),
                                   expression(d[pl]),
                                   expression(d[t])))+
  theme_classic(base_size=base_plot_size)
s5a

x <- readRDS("data/RESULTS/all_pars_wider_chain.Rds")[[1]]

par <- data.frame(x$par)
par <- par[round(nrow(par)/2):nrow(par),]
Nruns <- 500


ids <- sample(1:nrow(par),Nruns)


df <- lapply(1:Nruns, function(id) load_fit(as.character(id), par[id,]))
ll <- do.call(rbind,lapply(df, function(di) di$dfll))
pll <- ggplot(ll,aes(x=ll,y=llko))+
  geom_point(alpha=0.5)+
  scale_x_log10()+
  scale_y_log10()+
  geom_abline()
pll


par <- exp(par)
x <- par
par$id <- 1:nrow(par)

par <- reshape2::melt(par,id.vars="id")

levels(par$variable) <- c(k_LAG="k[l]",k_TIM="k[t]",k_PD="k[p]",k_inf="s[0]",
                          k_exp="s[e]",de="d[e]",d_lag="d[l]",d_tim="d[t]",
                          d_pd1="d[p]",d_pdl1="d[pl]",d_ifn="d[i]",
                          k_ifn="k[i]",ke="k[e]")[levels(par$variable)]
s4a <- ggplot(par,aes(x=value))+
  facet_wrap(~variable,scales="free",ncol = 5,labeller=label_parsed)+
  geom_histogram(bins = 20)+
  #geom_density()+
  scale_x_log10("parameter value",guide = guide_axis(check.overlap = TRUE))+
  scale_y_continuous("MCMC samples")+
  theme_classic(base_size=base_plot_size)
s4a

s5b <- ggplot(x,aes(x=k_LAG,y=d_lag))+
  #geom_point(alpha=0.1)+
  geom_bin_2d(show.legend=F)+
  scale_x_log10(expression(Lag3~inhibitor~strength~k[l]))+
  scale_y_log10(expression(Lag3~decay~rate~d[l]))+
  theme_classic()+
  scale_fill_viridis_c("MCMC\nsamples")
s5b
s5c <- ggplot(x,aes(x=k_TIM,y=d_tim))+
  #geom_point(alpha=0.1)+
  geom_bin_2d()+
  scale_x_log10(expression(Tim3~inhibitor~strength~k[t]))+
  scale_y_log10(expression(Tim3~decay~rate~d[t]))+
  theme_classic()+
  scale_fill_viridis_c("MCMC\nsamples")
s5c

df_inh <- do.call(rbind,lapply(df, function(di) di$df_inh))
df <- do.call(rbind,lapply(df, function(di) di$df))

d0 <- setup_opt(path = "data/PAR_EST/")
dat <- reshape_data(d0$dat)
dat1 <- dat[dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
dat <- dat[!dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
df$value[df$variable=="IFN"&df$value>3] <- NaN

id_renamr <- c(E="CTLs",g="g (day^-1)",gr="S-G2-M:G1",IFN="Ifng",
  LAG3="Lag3",TIM3="Tim3",PD1="PD1",PDL1="PDL1")
df$variable <- id_renamr[as.character(df$variable)]
dat$variable <- id_renamr[as.character(dat$variable)]
s5d <- ggplot(df)+
  facet_grid(rows=vars(variable),scales="free")+
  geom_line(aes(x=time,y=value,group=id),alpha=0.1)+
  geom_point(data=dat,aes(x=day,y=mean))+
  geom_errorbar(data=dat,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
  theme_classic(base_size=base_plot_size)
s5d

s5e <- ggplot(df_inh)+
  facet_grid(rows=vars(variable),scales="free")+
  geom_line(aes(x=time,y=value,group=id),alpha=0.1)+
  geom_point(data=dat1,aes(x=day,y=mean))+
  geom_errorbar(data=dat1,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
  theme_classic(base_size=base_plot_size)
s5e

top_row <- cowplot::plot_grid(s5a,labels="A")
middle_row <- cowplot::plot_grid(s5b,s5c,nrow=1,labels=c("B","C"),rel_widths = c(2.2,3))
bottom_row <- cowplot::plot_grid(s5d,s5e,nrow=1,labels=c("D","E"))

s5 <- cowplot::plot_grid(top_row,middle_row,bottom_row,nrow=3,rel_heights=c(3,3,4))

```
Study whether color coding s5E helps answer reviewer comment:

```{r}
base_plot_size=10
x <- readRDS("data/RESULTS/all_pars_wider_chain.Rds")[[1]]

par <- data.frame(x$par)
par <- par[round(nrow(par)/2):nrow(par),]

Nruns <- 100
ids <- sample(1:nrow(par),Nruns)
parsample <- par[ids[1:Nruns],]

df <- pbapply::pblapply(1:Nruns, function(id) load_fit(as.character(id), parsample[id,]))
ll <- do.call(rbind,lapply(df, function(di) di$dfll))
parsample <- exp(parsample)
x <- parsample
parsample$id <- 1:nrow(parsample)

df_inh <- do.call(rbind,lapply(df, function(di) di$df_inh))
df_inh$dp <- parsample$d_pd1[as.numeric(df_inh$id)]
df_inh$d_tim <- parsample$d_tim[as.numeric(df_inh$id)]
df_inh$d_lag <- parsample$d_lag[as.numeric(df_inh$id)]
df <- do.call(rbind,lapply(df, function(di) di$df))
df$dp <- parsample$d_pd1[as.numeric(df$id)]
df$d_tim <- parsample$d_tim[as.numeric(df$id)]
df$d_lag <- parsample$d_lag[as.numeric(df$id)]



d0 <- setup_opt(path = "data/PAR_EST/")
dat <- reshape_data(d0$dat)
dat1 <- dat[dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
dat <- dat[!dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
df$value[df$variable=="IFN"&df$value>3] <- NaN

id_renamr <- c(E="CTLs",g="g (day^-1)",gr="S-G2-M:G1",IFN="Ifng",
  LAG3="Lag3",TIM3="Tim3",PD1="PD1",PDL1="PDL1")
df$variable <- id_renamr[as.character(df$variable)]
dat$variable <- id_renamr[as.character(dat$variable)]

df$col.id <- "c1"
df_inh$col.id <- "c2"

df <- rbind(df,df_inh)



```


```{r}
alpha <- 0.6
df <- df[order(df$dp),]
datx <- rbind(dat,dat1)

df$var_f <- factor(df$variable,levels=c("CTLs", "Ifng", "g (day^-1)", "S-G2-M:G1", 
                                        "LAG3", "TIM3", "PD1","PDL1"))
datx$var_f <- factor(datx$variable,levels=c("CTLs", "Ifng", "g (day^-1)", "S-G2-M:G1", 
                                        "LAG3", "TIM3", "PD1","PDL1"))

s6a <- ggplot(df)+
  facet_wrap(~var_f,ncol=4,dir = "h",scales="free",labeller = label_parsed)+
  geom_line(aes(x=time,y=value,group=id,color=dp),alpha=alpha)+
  geom_point(data=datx,aes(x=day,y=mean))+
  geom_errorbar(data=datx,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
  theme_classic(base_size=base_plot_size)+
  scale_color_viridis_c(expression(d[p]),trans="log",breaks=c(0.06,0.4,3,20))
s6a


df <- df[order(df$d_lag),]

s6b <- ggplot(df)+
  facet_wrap(~var_f,ncol=4,dir = "h",scales="free",labeller = label_parsed)+
  geom_line(aes(x=time,y=value,group=id,color=d_lag),alpha=alpha)+
  geom_point(data=datx,aes(x=day,y=mean))+
  geom_errorbar(data=datx,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
  theme_classic(base_size=base_plot_size)+
  scale_color_viridis_c(expression(d[l]),trans="log",breaks=c(0.06,0.4,3,20))
s6b

df <- df[order(df$d_tim),]

s6c <- ggplot(df)+
  facet_wrap(~var_f,ncol=4,dir = "h",scales="free",labeller = label_parsed)+
  geom_line(aes(x=time,y=value,group=id,color=d_tim),alpha=alpha)+
  geom_point(data=datx,aes(x=day,y=mean))+
  geom_errorbar(data=datx,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
  theme_classic(base_size=base_plot_size)+
  scale_color_viridis_c(expression(d[t]),trans="log",breaks=c(0.06,0.4,3,20))
s6c

s6 <- cowplot::plot_grid(s6a,s6b,s6c,nrow=3,labels = c("A","B","C"))

```

Inhibitor behaviour study


```{r}
base_plot_size <- 12
ff <- c("ko_k_LAG_k_PD_opt_out.Rds","ko_k_LAG_k_TIM_opt_out.Rds",
        "ko_k_TIM_k_PD_opt_out.Rds","opt_out_wider_bounds.Rds",
        "ko_k_TIM_k_PD_k_LAG_opt_out.Rds")
ids <- c("Tim3","PD1/PDL1","Lag3","all","none")
fixpars <- list(c(k_LAG=0,k_PD=0),
                c(k_LAG=0,k_TIM=0),
                c(k_TIM=0,k_PD=0),
                NULL,
                c(k_TIM=0,k_PD=0,k_LAG=0))

best <- lapply(ff,function(ffi){
  x <- readRDS(paste0("data/RESULTS/",ffi))
  v <- sapply(x, function(xi) xi$value)
  
  x[[which.min(v)]]
})


best_fits <- lapply(1:length(best), function(i){
  id <- ids[i]
  pars <- best[[i]]$par
  fp <- fixpars[[i]]
  load_fit(id,pars,fp)
  
})

df <- do.call(rbind,lapply(best_fits,function(bi) bi$df))
df_inh <- do.call(rbind,lapply(best_fits,function(bi) bi$df_inh))

aic <- do.call(rbind,lapply(best,function(bi){
  npar <- length(bi$par)
  ll <- -bi$value
  aic <- 2*npar-2*ll
  data.frame(npar,ll,aic)
}))

aic$id <- ids

d0 <- setup_opt(path = "data/PAR_EST/")
dat <- reshape_data(d0$dat)
dat1 <- dat[dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
dat <- dat[!dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
id_renamr <- c(E="CTLs",g="g (day^-1)",gr="S-G2-M:G1",IFN="Ifng",
  LAG3="Lag3",TIM3="Tim3",PD1="PD-1",PDL1="PD-L1")

df$variable <- id_renamr[as.character(df$variable)]
dat$variable <- id_renamr[as.character(dat$variable)]

s3a <- ggplot(df)+
  facet_grid(rows=vars(variable),scales="free",labeller = label_parsed)+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id),show.legend = F)+
  geom_point(data=dat,aes(x=day,y=mean))+
  geom_errorbar(data=dat,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
    scale_x_continuous("days")+
  theme_classic()+
    scale_linetype_manual("",values=c("solid","dashed","solid","solid","dotted"))+
  scale_color_manual("",values=c("#000000","#006BFF","#07FF50","#B96BB2","#FF0505"))+
  scale_size_manual("",values=c(0.8,1.2,0.8,0.8,1.2))
s3a

df_inh$variable <- id_renamr[as.character(df_inh$variable)]
dat1$variable <- id_renamr[as.character(dat1$variable)]

s3b <- ggplot(df_inh)+
  facet_grid(rows=vars(variable),scales="free")+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id))+
  geom_point(data=dat1,aes(x=day,y=mean))+
  geom_errorbar(data=dat1,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
    scale_x_continuous("days")+
  scale_linetype_manual("",values=c("solid","dashed","solid","solid","dotted"))+
  scale_color_manual("",values=c("#000000","#006BFF","#07FF50","#B96BB2","#FF0505"))+
  scale_size_manual("",values=c(0.8,1,0.8,0.8,1))+
  theme_classic()
s3b

s3 <- cowplot::plot_grid(s3a,s3b,nrow=1,labels=c("A","B"),rel_widths = c(2.2,3))







```




```{r}


pars['d_pdl1'] <- pars['d_pdl1']+4


x1 <- load_fit("TIM\ndominant",pars)

df <- x1$df
df_inh <- x1$df_inh

d0 <- setup_opt(path = "data/PAR_EST/")
dat <- reshape_data(d0$dat)
dat1 <- dat[dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]
dat <- dat[!dat$variable%in%c("LAG3","TIM3","PD1","PDL1"),]


id_renamr <- c(E="CTLs",g="g (day^-1)",gr="S-G2-M:G1",IFN="Ifng",
  LAG3="Lag3",TIM3="Tim3",PD1="PD-1",PDL1="PD-L1")

df$variable <- id_renamr[as.character(df$variable)]
dat$variable <- id_renamr[as.character(dat$variable)]

p5b <- ggplot(df)+
  facet_grid(rows=vars(variable),scales="free",labeller = label_parsed)+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id),show.legend=F)+
  geom_point(data=dat,aes(x=day,y=mean))+
  geom_errorbar(data=dat,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
    scale_x_continuous("days")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(0.8,1))+
  theme_classic(base_size=base_plot_size)
p5b

df_inh$variable <- id_renamr[as.character(df_inh$variable)]
dat1$variable <- id_renamr[as.character(dat1$variable)]

p5c <- ggplot(df_inh)+
  facet_grid(rows=vars(variable),scales="free")+
  geom_line(aes(x=time,y=value,color=id,linetype=id,size=id))+
  geom_point(data=dat1,aes(x=day,y=mean))+
  geom_errorbar(data=dat1,aes(x=day,ymin=mean-sd,ymax=mean+sd))+
    scale_x_continuous("days")+
  scale_color_manual("",values=c("#006BFF","#FF0505"))+
  scale_linetype_manual("",values=c(5,3))+
  scale_size_manual("",values=c(0.8,1))+
  theme_classic(base_size=base_plot_size)
p5c

print(x1$dfll)
```

