---
title: "Interpolation model"
author: "Richard J Beck"
date: "11/25/2022"
output: html_document
---

This code generates the output for figure 4 and the associated supplementary figure.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="path/to/LAG_TIM_BIOMARKERS/")
```


```{r}
library(pbapply)
library(parallel)
library(deSolve)
library(DEoptim)
library(ggplot2)
#source("rev_01/Rscripts/ode_model.R")
source("code/utils/optim_funcs.R")
source("code/utils/plotting_funcs.R")
```
Phenomenological model: linearly interpolate CTL and IFN data
In this code, demonstrate the workings of the interpolation functions:
```{r}
require(ggplot2)

pdf <- readRDS("data/PAR_EST/fit_dat.rds")
Edat <- rbind(cbind(pdf$ctl_dat[,c("day","value")],id="ctl"),data.frame(day=c(-2,16),value=c(0,0),id=c("ctl")))
Edat <- Edat[order(Edat$day),]
IFNdat <- rbind(cbind(pdf$ifn_dat[,c("day","value")],id="ifn"),data.frame(day=c(-2,10,16),value=c(0,0,0),id=c("ifn")))
IFNdat <- IFNdat[order(IFNdat$day),]

dat <- rbind(Edat,IFNdat)

getE <- function(t){
  Edat <- data.frame(day=c(-2,1,3,5,7,10,16),value=c(0,0,10718,9667,4487,897,0)) ## boilerplate for convenience 
  x0 <- tail(Edat[Edat$day<t,],1)
  x1 <- head(Edat[Edat$day>=t,],1)
  
  dx <- (t-x0$day)/(x1$day-x0$day)
  y <- x0$value + dx*(x1$value - x0$value)
}

getIFN <- function(t){
  IFNdat <- data.frame(day=c(-2,1,3,5,7,10,16),value=c(0,0.04,1.52,0.29,0.06,0,0)) ## boilerplate for convenience 
  x0 <- tail(IFNdat[IFNdat$day<t,],1)
  x1 <- head(IFNdat[IFNdat$day>=t,],1)
  
  dx <- (t-x0$day)/(x1$day-x0$day)
  y <- x0$value + dx*(x1$value - x0$value)
}

x <- seq(-1,15,0.1)
y <- sapply(x,getE)
z <- sapply(x,getIFN)

p4a <- ggplot(subset(dat,id=="ctl"),aes(x=day,y=value))+
  geom_line(size=1.5)+
  theme_classic(base_size = 10)+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(CTLs~(mm^-3)))
p4a

p4b <- ggplot(subset(dat,id=="ifn"),aes(x=day,y=value))+
  geom_hline(yintercept = c(1.52,0.29,0.06),linetype=2,color=c(5,6,7))+
  geom_line(size=1.5)+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste(IFN,gamma," expression")))+
  theme_classic(base_size = 10)
p4b

```
Set eval sweep to true to actually repeat the sweep
```{r}
eval.sweep <- FALSE
source("code/utils/interpolation_optimiser.R")

```

Note this requires interpolation_optimiser.R script to be run to define the run.model function
```{r}

check_dec <- function(IFN){
  G1 <- 1/(1+8.5*IFN)
}

IFN <- seq(0,2,0.02)
dec <- sapply(IFN,check_dec)
dfdec <- data.frame(x=IFN,y=dec)


x<- readRDS("data/RESULTS/sweep_ke_ki.RDS")
x <- x[x$k_e<10,]
pdf <- readRDS("data/PAR_EST/fit_dat.rds")
pars <- x[x$err==min(x$err),c(1,2)]
pars2 <- c(pars$k_ifn,0)
pars3 <- c(0,pars$k_e)

parsdf <- rbind(pars,pars2,pars3)
parsdf$id <- c("max. likelihood", "no cytotoxicity", "no antiproliferative")

p4c <- ggplot(x,aes(x=k_e,y=k_ifn))+
  geom_raster(aes(fill=exp(-err)))+
  geom_point(data=parsdf,aes(shape=id),size=4)+
  theme_classic(base_size = 10)+
  scale_y_continuous(expression(antiproliferative~effect~(k[i])),breaks=seq(0,20,2))+
  scale_x_continuous(expression(cytotoxic~effect~(k[e])),breaks=seq(0,10,2))+
  scale_fill_viridis_c("likelihood",direction = 1)+
  scale_shape_discrete("")
p4c

p4cl <- ggplot(x,aes(x=k_e,y=k_ifn))+
  geom_raster(aes(fill=err))+
  geom_point(data=subset(parsdf,id=="max. likelihood"),aes(shape=id),size=4)+
  theme_classic(base_size = 10)+
  scale_y_continuous(expression(antiproliferative~effect~(k[i])),breaks=seq(0,20,2))+
  scale_x_continuous(expression(cytotoxic~effect~(k[e])),breaks=seq(0,10,2))+
  scale_fill_viridis_c("RMSE",direction = -1)+
  scale_shape_discrete("")
p4cl


run.mod <- function(pars){
  y <- run.model(pars)
  y$ratio <- y$SG2M/y$G1
  y$vol <- y$SG2M+y$G1
  modV <- y$vol
  y$growth <- c(log(modV[2:length(modV)]/modV[1:(length(modV)-1)])/diff(y$time),NA)## estimate growth rate between treatment points
  return(y)
}

y <- cbind(run.mod(pars),id="max. likelihood")
y2 <- cbind(run.mod(pars2),id="no cytotoxicity")
y3 <- cbind(run.mod(pars3),id="no antiproliferative")

y <- rbind(y,y2,y3)

p4e <- ggplot(dfdec,aes(x=x,y=y))+
  geom_vline(xintercept = c(1.52,0.29,0.06),linetype=2,color=c(5,6,7))+
  geom_line(size=1.5)+
  theme_classic(base_size=10)+
  scale_x_continuous(expression(paste(IFN,gamma," expression")))+
  scale_y_continuous("transition rate reduction",limits=c(0,1))
p4e

p4f <- ggplot(y,aes(x=time))+
  geom_line(size=1.5,aes(y=growth,color=id,linetype=id))+
  theme_classic(base_size=10)+
  geom_point(data=pdf$v_dat,aes(x=day,y=value))+
  geom_errorbar(data=pdf$v_dat,aes(x=day,ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(tumour~growth~rate~(day^-1)))+
  scale_color_discrete("")+
 scale_linetype_discrete("")+
  theme(legend.position = "none")
p4f

p4fleg <- ggplot(y,aes(x=time))+
  geom_line(size=1.5,aes(y=growth,color=id,linetype=id),show.legend = T)+
  theme_classic(base_size=10)+
  geom_point(data=pdf$v_dat,aes(x=day,y=value))+
  geom_errorbar(data=pdf$v_dat,aes(x=day,ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(tumour~growth~rate~(day^-1)))+
  scale_color_discrete("")+
 scale_linetype_discrete("")
p4fleg

p4fl <- ggplot(subset(y,id=="max. likelihood"),aes(x=time))+
  geom_line(size=1.5,aes(y=growth),show.legend = F)+
  theme_classic(base_size=10)+
  geom_point(data=pdf$v_dat,aes(x=day,y=value))+
  geom_errorbar(data=pdf$v_dat,aes(x=day,ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(tumour~growth~rate~(day^-1)))
p4fl



p4g <- ggplot(y,aes(x=time))+
  geom_line(size=1.5,aes(y=ratio,color=id,linetype=id))+
  geom_point(data=pdf$GR_dat,aes(x=day,y=value))+
  geom_errorbar(data=pdf$GR_dat,aes(x=day,ymin=value-sd,ymax=value+sd))+
  theme_classic(base_size=10)+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(S-G[2]-M~":"~G[1]~ratio))+
  scale_color_discrete("")+
 scale_linetype_discrete("")+
  theme(legend.position = "bottom")
p4g

p4gl <- ggplot(subset(y,id=="max. likelihood"),aes(x=time))+
  geom_line(size=1.5,aes(y=ratio))+
  geom_point(data=pdf$GR_dat,aes(x=day,y=value))+
  geom_errorbar(data=pdf$GR_dat,aes(x=day,ymin=value-sd,ymax=value+sd))+
  theme_classic(base_size=10)+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(S-G[2]-M~":"~G[1]~ratio))
p4gl

```


```{r}
pdf <- readRDS("data/PAR_EST/fit_dat.rds")
get_err_ke <- function(ke){
  pars <- c(0,ke)
  x <- run.model(pars)
  ssq(x,pdf)
}

get_err_kifn <- function(kifn){
  pars <- c(kifn,0)
  x <- run.model(pars)
  ssq(x,pdf)
}

get_err_full <- function(pars){
  x <- run.model(pars)
  ssq(x,pdf)
}

eval.sweep <- FALSE
source("code/utils/interpolation_optimiser.R")
opt0 <- optim(c(8,1),get_err_full)
opt1 <- optimise(get_err_ke,interval = c(0,20))
opt2 <- optimise(get_err_kifn,interval = c(0,20))

dfopt <- data.frame(rbind(c(opt0$par,-opt0$value),
               c(0,opt1$minimum,-opt1$objective),
               c(opt2$minimum,0,-opt2$objective)))
colnames(dfopt) = c("k_ifn","k_e","logL")
dfopt$aic <- c(3,2,2)*2-2*dfopt$logL
saveRDS(dfopt,"data/RESULTS/interpolated_fit_summary.rds")
```


```{r}

x<- readRDS("data/RESULTS/sweep_ke_ki.RDS")
x <- x[x$k_e<10,]
pdf <- readRDS("data/PAR_EST/fit_dat.rds")
dfopt <- readRDS("data/RESULTS/interpolated_fit_summary.rds")
pars <- dfopt[1,1:2]#x[x$err==min(x$err),c(1,2)]
pars2 <- dfopt[3,1:2] #c(pars$k_ifn,0)
pars3 <- dfopt[2,1:2]#c(0,pars$k_e)

parsdf <- rbind(pars,pars2,pars3)
parsdf$id <- c("max. likelihood", "no cytotoxicity", "no antiproliferative")

psa <- ggplot(x,aes(x=k_e,y=k_ifn))+
  geom_raster(aes(fill=exp(-err)))+
  geom_point(data=parsdf,aes(shape=id),size=4)+
  theme_classic(base_size = 10)+
  scale_y_continuous(expression(antiproliferative~effect~(k[i])),breaks=seq(0,20,2))+
  scale_x_continuous(expression(cytotoxic~effect~(k[e])),breaks=seq(0,10,2))+
  scale_fill_viridis_c("likelihood",direction = 1)+
  scale_shape_discrete("")
psa

run.mod <- function(pars){
  y <- run.model(pars)
  y$ratio <- y$SG2M/y$G1
  y$vol <- y$SG2M+y$G1
  modV <- y$vol
  y$growth <- c(log(modV[2:length(modV)]/modV[1:(length(modV)-1)])/diff(y$time),NA)## estimate growth rate between treatment points
  return(y)
}

y <- cbind(run.mod(pars),id="max. likelihood")
y2 <- cbind(run.mod(pars2),id="no cytotoxicity")
y3 <- cbind(run.mod(pars3),id="no antiproliferative")

y <- rbind(y,y2,y3)


psb <- ggplot(y,aes(x=time))+
  geom_line(size=1.5,aes(y=growth,color=id,linetype=id))+
  theme_classic(base_size=10)+
  geom_point(data=pdf$v_dat,aes(x=day,y=value))+
  geom_errorbar(data=pdf$v_dat,aes(x=day,ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(growth~rate~(day^-1)))+
  scale_color_discrete("")+
 scale_linetype_discrete("")+
  theme(legend.position = "none")
psb

psc <- ggplot(y,aes(x=time))+
  geom_line(size=1.5,aes(y=ratio,color=id,linetype=id))+
  geom_point(data=pdf$GR_dat,aes(x=day,y=value))+
  geom_errorbar(data=pdf$GR_dat,aes(x=day,ymin=value-sd,ymax=value+sd))+
  theme_classic(base_size=10)+
  scale_x_continuous("days after CTL transfer",breaks=seq(0,15,3))+
  scale_y_continuous(expression(S-G[2]-M~":"~G[1]~ratio))+
  scale_color_discrete("")+
 scale_linetype_discrete("")
psc


```